Extrinsic Type Inference (DRAFT 2025-06-08)

### 2025-07-09 extrinsic_type_inference_paper/paper.tex
add comment on bidirectional type inference

\TODO{add requirements on packing, polar sub and tidying, and negatble checks}

\TODO{for polar sub, dye type by subbing in a fresh skolem}

@@

\TODO{mention impredicative types with subtyping, bidirectional typing, local type inferenc (Pierce)}

\TODO{mention how inferring parameter types and instations is processed before and independent from their subtyping check}

@@

\section{Conclusion}
\label{sec:conclusion}
We have presented a novel extrinsic type inference system
that can guarantee programs are bug-free while maintaining
code-reusability nearly on par with untyped scripting languages,
and exceeding that of an previous type inference system we know of.  
We have incorporated previous studied ideas of subtyping constraints,
intersection types, union types, parametric types, and recursive types into
a cohesive design that enables more principal types to be inferred 
and more subtyping relations to be verified than in previous systems. 
In particular, our system is able to learn recursive types of function inputs
and outputs purely from the construction of recursive functions without 
any intrinsic recursive type specifications.
Additionally, our system can perform inductive reasoning to
verify than one recursive type subtypes another.
In the future, we will extend the recursive type reasoning capabilities
to handle recursive relational types, enabling the infinite paths
of recursive functions to be represented and reasoned with at the type level. 
With inductive reasoning over relational types, we will be
moving closer to a full fledged theorem proving system,
in which subtyping enables a form of proof automation,
by reusing the proof of one proposition as a proof for another.

>>>>

\TODO{add requirements on factoring, packing, polar sub and tidying, and negatble checks}

@@

To reduce the annotation burden and increase expressivity, researchers
developed ways to infer type instantiations and compare type with varying
amounts of possible instantiations, by combining impredicative types
with subtyping and type inference techniques. Bidirectional type inference \cite{}  
is one such technique where bounds on type variables can be inferred in 
both directions of the control flow.
It propagates inferred and annotated knowledge directly from
the structures of the program and relies on subtyping merely to check compatibility between
types. Additionally, whether a non-parametric type is an instantiation
of a parametric type is treated as distinct from subtyping.

@@

\section{Conclusion}
\label{sec:conclusion}
We have presented a novel extrinsic type inference system
that can guarantee programs are bug-free while maintaining
code-reusability nearly on par with untyped scripting languages,
and exceeding that of an previous type inference system we know of.  
We have incorporated previous studied ideas of subtyping constraints,
intersection types, union types, parametric types, and recursive types into
a cohesive design that enables more principal types to be inferred 
and more subtyping relations to be verified than in previous systems. 
In particular, our system is able to learn recursive types of function inputs
and outputs purely from the construction of recursive functions without 
any intrinsic recursive type specifications.
Additionally, our system can perform inductive reasoning to
verify than one recursive type subtypes another.
In the future, we will extend the recursive type reasoning capabilities
to handle recursive relational types, enabling the infinite paths
of recursive functions to be represented and reasoned with at the type level. 
With inductive reasoning over relational types, we will be
moving closer to a full fledged theorem proving system,
in which subtyping enables a form of proof automation,
by reusing the proof of one proposition as a proof for another. 
We may also explore different strategies for rewriting types 
to get as much expressivity as possible.
